---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

# dmpredict

## Overview

dmpredict is a package written to implement the prediction equations reported in `NCD Risk Factor Collaboration (2023)` to estimate the probability that an individual had HbA1c ≥6.5% based on measured fasting plasma glucose (FPG), and vice versa. The details of the statistical methods are provided in the citation below. In short, the prediction equations are generalised linear mixed models fitted with package `lme4`. The model objects in this package provide draws of the model coefficients to allow estimation of the uncertainty associated with the prediction equations.

<span style = "background-color:#f3f3f3">NCD Risk Factor Collaboration (NCD-RisC). Global variation in diabetes diagnosis and prevalence based on fasting glucose and hemoglobin A1c. *Nature Medicine* 2023; **29**: 2885-2901.</span>

The users must provide  at least one of the following:

* `fpg_clean` cleaned FPG (mmol/L).
* `hba1c_clean` cleaned HbA1c (%).

Plus the following required predictors:

* `sex` coded as "female" and "male".
* `age` in years (current models have only been tested in adults aged 18+ years).
* `study_id` unique identifier for each study (or study round).
* `iso` [ISO-3 alpha code](https://en.wikipedia.org/wiki/ISO_3166-1_alpha-3) for the country the study was conducted in.
* `medication` whether using medication for diabetes, coded as 1 = yes, 0 = no, NA = missing.

Optional covariates for prediction:

* `bmi_clean` cleaned body-mass index (kg/m^2^).
* `handheld_fpg` how FPG was measured, coded as 1 = using a POC handheld device, 0 = lab based measurement, NA = unknown.
* `handheld_a1c` how HbA1c was measured, coded as 1 = using a POC handheld device, 0 = lab based measurement, NA = unknown.

If `iso` is not recognized, provide a string with one of the following in a variable named `Superregion`:

* "Central and Eastern Europe"
* "Central Asia, Middle East and North Africa"
* "High-income western"
* "Latin America and Caribbean"
* "Oceania"
* "Southeast Asia, East Asia and the Pacific"
* "South Asia"
* "Sub-Saharan Africa"

## Installation

To install the package, use:

`remotes::install_github("NCD-RisC/dmpredict")`

## Usage

```{r, eval = FALSE}
library(dmpredict)

# Make sure 'data' object has the minimum columns as stated above

df <- transform_data(data)

# Predict_missing_fpg or hba1c functions return a probability that the predicted biomarker
# is over the diagnostic threshold for diabetes, i.e. FPG >=7.0 mmol/L and HbA1c >=6.5%

data$predicted_a1c <- predict_missing_hba1c(df, cw_model_coefs, return_draws = FALSE, verbose = FALSE)
data$predicted_fpg <- predict_missing_fpg(df, cw_model_coefs, return_draws = FALSE, verbose = FALSE)

# using 'return_draws = TRUE' will return the N draws (default = 2000) of predictions in a matrix

res1 <- predict_missing_hba1c(df, cw_model_coefs, return_draws = TRUE, verbose = FALSE)
res2 <- predict_missing_fpg(df, cw_model_coefs, return_draws = TRUE, verbose = FALSE)

data$predicted_a1c2 <- res1$predicted_a1c
data$predicted_fpg2 <- res2$predicted_fpg
print(dim(res1$mat_predicted_a1c))
print(dim(res2$mat_predicted_fpg))

# using a different number of simulation draws

res3 <- predict_missing_hba1c(df, cw_model_coefs, N_sim = 5000, return_draws = TRUE, verbose = FALSE)
res4 <- predict_missing_fpg(df, cw_model_coefs, N_sim = 5000, return_draws = TRUE, verbose = FALSE)

print(dim(res3$mat_predicted_a1c))
print(dim(res4$mat_predicted_fpg))

```
